Collaborative Test-Driven Scientific Model Validation

### 2013-11-23 sciunit-icse14/sciunit-icse14.tex
First few paragraphs of new intro

\begin{abstract}
One of the pillars of the modern scientific method is \emph{model validation}: comparing a scientific model's predictions against empirical observations. Today, a scientist demonstrates the validity of a model by making an argument in a paper and submitting it for peer review, a process comparable to \emph{code review} in software engineering. While peer review ensures that contributions meet high-level goals, software engineers typically supplement this process with \emph{unit testing} to get a more complete picture of the state of a software project, particularly for complex projects involving many developers.

We argue that a similar test-driven methodology would be valuable to scientific communities that need to validate increasingly complex models against growing collections of empirical data. The dynamics of scientific communities differ from software communities in several key ways, however. In this paper, we introduce \emph{SciUnit}, a test-driven model validation framework designed to handle these differences in a practical manner. We also describe how SciUnit, supported by both new and existing collaborative infrastructure, can be integrated into the modern scientific process.
%in the context of the applications and architectures that the language may need to target.
%
%Programming languages targeted toward end-users in high-performance computing must consider a number of issues in addition to performance. Typically, researchers have focused verifiability, portability and ease-of-use, and the right balance often depends on the particular uses that they are put toward.
%
%Ace %, evaluates it against several design and adoption criteria, and describes several use cases as well as a realistic case study involving parallelizing neurobiological circuit simulations on clusters of GPUs. 
%is an extensible statically-typed programming language embedded within the ubiquitous scripting language Python. Python serves as a {metalanguage}, enabling programmatic control over several aspects of language specification and compilation. 
%We demonstrate the power of this design for HPC by defining all of the primitives of the OpenCL kernel language as a library of user-defined types in Ace, then extending this core with several higher-level first-class abstractions. The target of translation, called the \emph{backend}, is user-defined as well, allowing us to implement the semantics of these primitives by direct translation to OpenCL, CUDA or C99.
%
%Although Ace is built around a static type discipline, it supports a novel form of generic programming based on type propagation that, together with a form of extensible local type inference, eliminates most of the syntactic overhead usually associated with statically-typed languages. A companion host API allows type information from \verb|numpy| arrays to directly propagate into Ace kernels, supporting seamless adoption of Ace into current scientific workflows.
%
\end{abstract}

@@

\section{Introduction}

Neuroscientists construct quantitative models to coherently explain observations of neurons, circuits, brain regions and behavior. 
These models can be characterized by their \textit{scope}: the set of observable quantities that the model can generate predictions about, and by their \textit{validity}: the extent to which these predictions agree with empirical observations of those quantities.

Today, scientists contribute a new model to the research community by submitting a paper describing how the model works, along with selected figures demonstrating its scope and validity and comparing its validity to other models with the same scope. 
Reviewers are then responsible for discovering relevant data and competing models that the paper did not adequately consider, drawing on their knowledge of prior publications. 
However, in many areas, the number of publications being generated every year can overwhelm even the most conscientious scientists \cite{jinha_article_2010}.

>>>>

\begin{abstract}
One of the pillars of the modern scientific method is \emph{model validation}: comparing a scientific model's predictions against empirical observations. Today, a scientist demonstrates the validity of a model by making an argument in a paper and submitting it for peer review, a process comparable to \emph{code review} in software engineering. While human review helps to ensure that contributions meet high-level goals, software engineers typically supplement this with \emph{unit testing} to get a more complete picture of the status of a software project, particularly for complex projects involving many developers.

We argue that a similar test-driven methodology would be valuable to scientific communities as they seek to  validate increasingly complex models against growing collections of empirical data. The dynamics of  scientific communities and software communities differ in several key ways, however. In this paper, we introduce \emph{SciUnit}, a framework for test-driven scientific model validation. We describe how SciUnit, supported by new and existing collaborative infrastructure, can integrate into the modern scientific process.
%in the context of the applications and architectures that the language may need to target.
%
%Programming languages targeted toward end-users in high-performance computing must consider a number of issues in addition to performance. Typically, researchers have focused verifiability, portability and ease-of-use, and the right balance often depends on the particular uses that they are put toward.
%
%Ace %, evaluates it against several design and adoption criteria, and describes several use cases as well as a realistic case study involving parallelizing neurobiological circuit simulations on clusters of GPUs. 
%is an extensible statically-typed programming language embedded within the ubiquitous scripting language Python. Python serves as a {metalanguage}, enabling programmatic control over several aspects of language specification and compilation. 
%We demonstrate the power of this design for HPC by defining all of the primitives of the OpenCL kernel language as a library of user-defined types in Ace, then extending this core with several higher-level first-class abstractions. The target of translation, called the \emph{backend}, is user-defined as well, allowing us to implement the semantics of these primitives by direct translation to OpenCL, CUDA or C99.
%
%Although Ace is built around a static type discipline, it supports a novel form of generic programming based on type propagation that, together with a form of extensible local type inference, eliminates most of the syntactic overhead usually associated with statically-typed languages. A companion host API allows type information from \verb|numpy| arrays to directly propagate into Ace kernels, supporting seamless adoption of Ace into current scientific workflows.
%
\end{abstract}

@@

\section{Introduction}
Scientific theories often take the form of a \emph{quantitative model}: a formal structure that can generate predictions about observable quantities. Such a model is characterized by its \textit{scope}: the set of observable quantities that the model can predict, and by its \textit{validity}: the extent to which its predictions agree with experimental observations of these quantities.

Quantitative models are described and validated using a social process: the scientific publication system. For a model to be accepted by the scientific community, scientists must write a paper describing it and providing  evidence that it predicts some quantity of interest more accurately than previous models, or that the tradeoff it makes between accuracy and complexity may be useful to the community \cite{box}. During the \emph{peer review} process, other members of the relevant community are tasked with ensuring that the validation criteria used (e.g. a $p$-value) are strong enough and that all relevant data and competing models were fully considered, drawing on knowledge of statistical methods and the prior literature. Publishing a paper is one of the primary motivators for most scientists \cite{howison2011scientific}.

Quantitative models have much in common with software. Indeed, quantitative models are increasingly  implemented as computer programs. In some cases the program implements a mathematical model, while in other cases the program \emph{is} the model (e.g. complex simulations). The peer review process is similar in many ways to the \emph{code review} process used by many development teams \cite{codereview}. During code review, team members look for errors, enforce style and architectural guidelines and check that the code is valid (i.e. that achieves its intended goal \cite{validity}) before it is committed to a primary source code repository. 
Code review is reasonably effective \cite{creffectiveness1,creffectiveness2}. However, to be most effective reviewers must expend considerable effort \cite{effort}, and code review is most effective for resolving  issues related to software architecture \cite{evolvability}.

For these reasons, software development teams typically supplement code reviews with a number of more \emph{automated} approaches. One of the most widely-adopted approaches is \emph{unit testing}.

--------

However, in many areas, the number of publications being generated every year can overwhelm even the most conscientious scientists \cite{jinha_article_2010}.
